---
title: "Regression_Models_Week3"
author: "Dimira Petrova"
date: "April 11, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as confounder. Give the adjusted estimate for the expected change in mpg comparing 8 cylinders to 4.

```{r}
# This model is adjusted to the weight
data(mtcars)
fit = lm(mpg ~ factor(cyl) + wt, mtcars)
summary(fit)$coef
```

Answer: -6.070860

## Question 2

Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight as a possible confounding variable. Compare the effect of 8 versus 4 cylinders on mpg for the adjusted and unadjusted by weight models. Here, adjusted means including the weight variable as a term in the regression model and unadjusted means the model without weight included. What can be said about the effect comparing 8 and 4 cylinders after looking at models with and without weight included?.

```{r}
# Fit a model that does not contain wt as a confounder or term.
# This is the unadjusted weight models.
fit2 = lm(mpg ~ factor(cyl), data = mtcars)
# Compare the estimate of 8 cyls compared to that of 4 cyls from the models with and without wt.
cyl.8 = data.frame(summary(fit)$coef[3], summary(fit2)$coef[3]); names(cyl.8) = c("with wt", "no wt"); rownames(cyl.8) = "cyl 8 Est."
print(cyl.8)
```
Answer: Holding weight constant, cylinder appears to have less of an impact on mpg than if weight is disregarded.

## Question 3

Consider the mtcars data set. Fit a model with mpg as the outcome that considers number of cylinders as a factor variable and weight as confounder. Now fit a second model with mpg as the outcome model that considers the interaction between number of cylinders (as a factor variable) and weight. Give the P-value for the likelihood ratio test comparing the two models and suggest a model using 0.05 as a type I error rate significance benchmark.

```{r}
# Fit a model with mpg as the outcome with cyl as a factor variable and weight as confounder.
fit3 = lm(mpg ~ factor(cyl) + wt, data = mtcars)
# Fit a model with mpg as the outcome that considers the interaction between number of cylinders and weight.
fit4 = lm(mpg ~ factor(cyl)*wt, data = mtcars)
anova(fit3, fit4)
```

Answer: The P-value is larger than 0.05. So, according to our criterion, we would fail to reject, which suggests that the interaction terms may not be necessary.

## Question 4
Consider the mtcars data set. Fit a model with mpg as the outcome that includes number of cylinders as a factor variable and weight inlcuded in the model as
```{r}
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
```
How is the wt coefficient interpretted?
```{r}
lm(mpg ~ I(wt * 0.5) + factor(cyl), data = mtcars)
```
Answer: The estimated expected change in MPG per one ton increase in weight for a specific number of cylinders(4, 6, 8).

## Question 5
Consider the following data set
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```
Give the hat diagonal for the most influential point
```{r}
x = c(0.586, 0.166, -0.042, -0.614, 11.72)
y = c(0.549, -0.026, -0.127, -0.751, 1.344)
fit5 = lm(y ~ x)
# Find the maximum point.
hatvalues(fit5)[which.max(abs(hatvalues(fit5)))]

# or it can be done by
influence(fit5)$hat

## showing how it's actually calculated
xm <- cbind(1, x)
diag(xm %*% solve(t(xm) %*% xm) %*% t(xm))

```
Answer: 0.9946

## Question 6
Consider the following data set
```{r}
x <- c(0.586, 0.166, -0.042, -0.614, 11.72)
y <- c(0.549, -0.026, -0.127, -0.751, 1.344)
```
Give the slope dfbeta for the point with the highest hat value.
```{r}
influence.measures(fit5)$infmat[which.max(abs(influence.measures(fit5)$infmat[, 2])), 2]
```
Answer: -133.8226

## Question 7
Consider a regression relationship between Y and X with and without adjustment for a third variable Z. Which of the following is true about comparing the regression coefficient between Y and X with and without adjustment for Z.

Answer: It is possible for the coefficient to reverse sign after adjustment. For example, it can be strongly significant and positive before adjustment and strongly significant and negative after adjustment.